---
title: "GAM evaluation - SF Bay"
output: 
  html_document:
    code_folding: hide
    css: kable.css
self_contained: yes
runtime: shiny
---
<a href="https://github.com/fawda123/SFBaytrends/" class="github-corner" aria-label="View source on Github"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#FD6C6C; color:#fff; position: absolute; top: 0; border: 0; left: 0; transform: scale(-1, 1);" aria-hidden="true"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a><style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style>

```{r, message = F, warning = F, results = 'hide'}
# globals
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(lubridate)
library(mgcv)
library(shiny)
library(plotly)

source('R/funcs.R')

data(datprc)
data(modssta)
```

## Exploratory plots

```{r}
# create plots for whole time series, by year, by month
# all variables, stations
rawplo <- datprc %>% 
  group_by(param) %>% 
  nest %>% 
  mutate(
    tot = map(data, function(x){
      
      p <- ggplot(x, aes(x = dec_time, y = value)) + 
        geom_line() +
        theme_bw(base_family = 'serif') +
        facet_wrap(~ station, ncol = 2) +
        theme(strip.background = element_blank())
      
      return(p)
      
    }),
    yrs = map(data, function(x){
      
      p <- ggplot(x, aes(x = factor(yr), y = value)) + 
        geom_boxplot(base_family = 'serif') +
        theme_bw(base_family = 'serif') +
        scale_x_discrete("Year") +
        facet_wrap(~ station, ncol = 2) +
        theme(strip.background = element_blank())
      
      return(p)
      
    }), 
    mos = map(data, function(x){
      
      p <- ggplot(x, aes(x = mo, y = value)) + 
        geom_boxplot() +
        theme_bw() +
        facet_wrap(~ station, ncol = 2) +
        theme(strip.background = element_blank())
      
      return(p)
      
    })
  )
```

Plots for raw data, all stations and parameters, 1990 - 2017.  Select the parameter, plot type (total time series, by year, or by month), and variable transformation.
```{r, echo = F}
# selection widgets
selectInput("prm", "Choose parameter:", unique(datprc$param))
selectInput("typ", "Choose plot type:", c('tot', 'yrs', 'mos'))
selectInput("log", "Log-space:", c(T, F))
```
```{r}
# get the plot
toplo <- reactive({

  sel <- filter(rawplo, param %in% input$prm) %>% 
    select(matches(input$typ)) %>% 
    .[[1]] %>% 
    .[[1]]
  
  if(input$log){
  
    sel <- sel + scale_y_log10(input$prm)
    
  } else {
    
    sel <- sel + scale_y_continuous(input$prm)
    
  }
  
  return(sel)
  
})

# make the plot
renderPlot({
  
  toplo()
  
}, height = 800, width = 900)

```

## GAMs of log-chlorophyll with annual, seasonal trends

```{r}
# get GAMs for each station, gam0, gam1, and gam2 from baytrends

# smooths to evaluate
frms <- c(
  "log10(chl) ~ dec_time + s(doy, bs = 'cc')",  
  "log10(chl) ~ dec_time + s(dec_time) + s(doy, bs = 'cc')",
  "log10(chl) ~ dec_time + s(dec_time) + s(doy, bs = 'cc') + ti(dec_time, doy, bs = c('tp', 'cc'))"
  ) %>% 
  as.list

# enframe frms for combine with tomod
frms <- frms %>% 
  enframe('modi', 'frm')

# data to model, same as datprc, params in wide format, nested by station
# crossed with frms
tomod <- datprc %>% 
  spread(param, value) %>% 
  group_by(station) %>% 
  nest %>% 
  crossing(frms)

# # create models for every station, gam model eval
# modssta <- tomod %>%
#   mutate(
#     modv = pmap(list(data, frm), function(data, frm){
# 
#       # get upper bounds of knots
#       gamk1 <- data$yr %>%
#         unique %>%
#         length
#       gamk1 <- round(gamk1 * 2 / 3, 0) %>% 
#         pmax(10, .)
# 
#       # insert upper gamk1 if applicable
#       if(grepl('s\\(dec_time', frm)){
#         p1 <- gsub('(^.*)s\\(dec\\_time\\).*$', '\\1', frm)
#         p3 <-  gsub('^.*s\\(dec\\_time\\)(.*)$', '\\1', frm)
#         p2 <- paste0('s(dec_time, k = ', gamk1, ')') 
#         frm <- paste0(p1, p2, p3)
#       }
#       
#       gam(as.formula(frm),
#         knots = list(doy = c(1, 366)),
#         data = data,
#         na.action = na.exclude,
#         select = T
#       )
# 
#   })
#   )
# modssta <- modssta %>% 
#   mutate(modi = factor(modi, levels = c(1, 2, 3), labels = c('gam0', 'gam1', 'gam2')))
# save(modssta, file = 'data/modssta.RData', compress = 'xz')
```

Select station for summary of all three GAM chlorophyll models:
```{r, echo = F}
# selection widgets
selectInput("sta", "Choose station:", unique(datprc$station))
```
```{r}
mods <- reactive({
  modssta %>% filter(station %in% input$sta)
})

# smoother stats of GAMs
renderTable({
  map(mods()$modv, ~ summary(.x)$s.table %>% data.frame %>% rownames_to_column('smoother')) %>% 
    enframe %>% 
    rename(model = name) %>% 
    mutate(model = factor(model, levels = c(1, 2, 3), labels = c('gam0', 'gam1', 'gam2'))) %>% 
    unnest
})

# summary stats of GAMs, model fit/perf
renderTable({
  map(mods()$modv, ~ data.frame(
      AIC = AIC(.x), 
      GCV = .x$gcv.ubre,
      R2 = summary(.x)$r.sq)) %>% 
    enframe %>% 
    rename(model = name) %>% 
    mutate(model = factor(model, levels = c(1, 2, 3), labels = c('gam0', 'gam1', 'gam2'))) %>% 
    unnest
})
```

```{r}
# prediction data
prdplo <- reactive({
    
  prdplo <- mods() %>%
    mutate(prddat = pmap(list(data, modv), function(data, modv){
  
  
      prddat <- data.frame(
          dec_time = seq(min(data$dec_time), max(data$dec_time), length = 1000)
        ) %>%
        mutate(
          date = date_decimal(dec_time),
          date = as.Date(date),
          mo = month(date, label = TRUE),
          doy = yday(date),
          yr = year(date)
        )
  
      prd <- predict(modv, newdata = prddat)
      
      prddat <- prddat %>% mutate(chl = prd)
  
      return(prddat)
      
    })) %>% 
    select(modi, prddat) %>% 
    unnest
  
})

rawchl <- reactive({
  rawchl <- mods()$data[[1]]
})

# facet by mod, doy x axis, lines by year
renderPlot({
  p <- ggplot(prdplo(), aes(x = doy, group = factor(yr), colour = yr)) + 
    geom_line(aes(y = chl)) + 
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank()
      ) + 
    scale_y_continuous('log10(chl)') +
    facet_wrap(~ modi, ncol = 3) +
    guides(colour = guide_colourbar(barheight = 1, barwidth = 20)) + 
    ggtitle(paste0('Station', input$sta))
  p
}, height = 500, width = 900)

# plot predictions, all mods, by time
renderPlot({
  # plot
  p <- ggplot(prdplo(), aes(x = date)) + 
    geom_point(data = rawchl(), aes(y = log10(chl)), size = 0.5) +
    geom_line(aes(y = chl, colour = factor(modi))) + 
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank()
      ) + 
    ggtitle(paste0('Station', input$sta))
  p
})
```

### Individual model for selected station:

```{r, echo = F}
selectInput('mod', "Select a model:", unique(modssta$modi))
```
```{r}
# smoother stats of GAMs
renderTable({
  
  md <- mods() %>% 
    filter(modi %in% input$mod) %>% 
    .$modv %>% 
    .[[1]]
  tab <- summary(md)$s.table %>% data.frame %>% rownames_to_column('smoother')
  return(tab)

})

# summary stats of GAMs, model fit/perf
renderTable({
  
  md <- mods() %>% 
    filter(modi %in% input$mod) %>% 
    .$modv %>% 
    .[[1]] 
  tab <- data.frame(
    AIC = AIC(md), 
    GCV = md$gcv.ubre,
    R2 = summary(md)$r.sq)
  return(tab)
  
})
renderPlot({
  
  toplo <- prdplo() %>% 
    filter(modi %in% input$mod)
  
  # plot
  p <- ggplot(toplo, aes(x = date)) + 
    geom_point(data = rawchl(), aes(y = log10(chl)), size = 0.5) +
    geom_line(aes(y = chl)) + 
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank()
      ) + 
    ggtitle(paste0('Station', input$sta))
  p
})
renderPlot({
  md <- mods() %>% 
    filter(modi %in% input$mod) %>% 
    .$modv %>% 
    .[[1]] 
  gam.check(md)
})
```

```{r}
# prediction data
prdplost <- reactive({
  
  mods() %>% 
    filter(modi %in% input$mod) %>%
    mutate(prddat = pmap(list(data, modv), function(data, modv){

      data <- data$date %>% 
        range(na.rm  = T)
      data <- seq.Date(data[1], data[2], by = 'days') %>% 
        tibble(date = .) %>% 
        mutate(
          doy = yday(date), 
          dec_time = decimal_date(date),
          yr = year(date)
          )
      prd <- predict(modv, newdata = data)
      out <- data.frame(data, chl = prd)
      
      return(out)
      
    })
  
  ) %>% 
  select(modi, prddat) %>% 
  unnest %>% 
  select(-modi)
  
})
```

```{r, out.width = "100%"}
renderPlotly({
  
  tosrf <- prdplost() %>% 
    select(-date, -dec_time) %>% 
    spread(yr, chl) %>% 
    select(-doy) %>% 
    as.matrix
  
  scene <- list(
    aspectmode = 'manual', 
    aspectratio = list(x = 1, y = 1, z = 0.5), 
    xaxis = list(title = 'Years from time zero'), 
    yaxis = list(title = 'Day of year'), 
    zaxis = list(title = 'log-Chl'),
    roughness = .01,
    ambient = 0.9
  )
  
  p <- plot_ly(z = ~tosrf) %>% 
    add_surface(colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
    layout(scene = scene)
  p
  
})
